\chapter{Conclusions}
\indent

The purpose of this work was to understand nitrogen doped chemical vapor deposition of diamond such that optimized conditions could be selected for any desired outcome. In this case, the desired outcome is the ideal substrate for laser writing of nitrogen vacancy (NV) defect arrays which consists of a high quality single crystal diamond plate with a delta doped layer of single substitutional nitrogen. With an ideal substrate, spatial control of NV centers via laser writing will be significantly improved thus improving the spatial resolution of quantum sensors. Determining optimized conditions included thorough understanding of substrate surface preparation on deposition quality, established growth quality as growth rate significantly decreased, and a parametric study of reactor parameter effects on nitrogen incorporation. A summary of each investigation and concluding remarks on the success and proposed future direction of each work will be presented here.

\section{Substrate Surface Preparation}
\indent

Several methods for substrate surface preparation were evaluated in Chapter \ref{chSurfacePrep}. First, every substrate was required to have surface roughness of not more than 1 nm after mechanical polishing. Then, an extended hydrogen plasma etch was incrementally administered to a single substrate. Microscale surface features and nanoscale surface roughness was evaluated after each increment. The hydrogen plasma etch preferentially etched polishing damage and dislocations in the substrate. This increased the surface roughness of the sample, after an slight decrease in the first 15 minute etch increment, as etch time progressed. The same experiment was performed on a second substrate with few etch steps. Again, the surface roughness increased as cumulative etch time increased. The increase in surface roughness was a result of increased etch pits with exposed crystalline facets. The removal of amorphous polishing damage and exposure of crystalline facets resulted in low strain grown films. Lastly, a reactive ion etch was performed on the same starting substrate as the first hydrogen etch experiment, however, the grown film was removed and the surface was mechanically polished to not more than 1 nm surface roughness. The reactive ion etch ballistically removed 6.5 $\mu m$ of polishing damaged material from the center of the diamond substrate exposing clean single crystal diamond. The reactive ion etch preferentially etch the polishing lines, but exposed a smooth, low roughness, diamond surface. The reactive ion etch also produced a low strain high quality single crystal diamond film. The extended hydrogen plasma etch was implemented as the surface normalization process due to its ease of access during chemical vapor deposition of diamond, however, reactive ion etching demonstrated equivalent success in surface normalization.

\section{Low Methane Depositions}
\indent

Manufacturing of layered structures through CVD growth potentially requires the ability to slow down or halt growth rates during the deposition for changes in feed gas composition. The reactor used for this was was optimized for high growth rates of large single crystal diamond wafers. An experimental deposition series was completed to ensure a high quality of growth outside of the optimized growth regime of the reactor. Three samples at 3\%, 2\%, and 1\% methane with no intentional nitrogen doping were grown. Step flow growth, an indicator of high quality growth, was observed on all three depositions. Unepitaxial crystallites formed on the 1\% methane deposition indicating faster growth propagation from nucleation sites than overall vertical growth rate. In other words, the overall vertical plane growth rate was did not outpace the growth of adatom nucleation site from early in the deposition process.

In addition to growth morphology, NV background was measured using photoluminescent spectroscopy. Each spectra was normalized to the Raman line allowing for qualitative comparison of intensities. As methane concentration decreased, NV concentration increased. This is a product of a constant leak rate of the reactor at constant pressure allowing atmospheric nitrogen in the deposition chamber. A constant flow rate of nitrogen  while methane concentration is decreased creates a higher ratio of nitrogen to carbon allowing more nitrogen to incorporate in the single crystal diamond growth.

The high NV background at reduced growth rates defeats the purpose utilizing slow growth rates to exchange feed gas composition and improve layered structures. The high growth rate reactor is inadequate for layered structure growths due to large residence times and atmospheric nitrogen leak rate. Post processing, such as high pressure annealing and exposure to a high temperature hydrogen environment, did not reduce NV background of low methane depositions to a usable level for laser writing. 

\section{Effects of Pressure and Offcut on Nitrogen Incorporation}
\indent

Within the layered structure, substitutional nitrogen is the ideal nitrogen defect for laser writing arrays of single quantum sensors. To preferentially incorporate nitrogen as substitutional nitrogen, a parametric study of reactor parameters was conducted to identify deposition conditions to suppress the incorporation of nitrogen as NV$^-$ and favor the incorporation of nitrogen as substitutional nitrogen. Deposition parameters evaluated include substrate temperature, reactor pressure, and substrate offcut. The first experimental series emphasized the a correlation in nitrogen incorporation and the confounding of variable surface roughness.

The confounding variable of surface roughness was mitigated through a surface normalization procedure developed in Chapter \ref{chSurfacePrep}. The surface normalization procedure consisted of a mechanical polish of not more than 1 nm surface roughness followed by a 5 hour hydrogen plasma etch immediately prior to deposition. The second experimental series shown no correlation between nitrogen related defect incorporation and surface roughness was observed indicated an effective surface normalization procedure.

Using the second experimental series with normalized surface condition, the effects of reactor pressure and substrate offcut on nitrogen incorporation was investigate. A high nitrogen concentration in the gas phase of the deposition was used to exaggerate the effects. At high nitrogen concentrations, the sensitivity of reactor pressure and substrate surface orientation was too low to favor the incorporation of nitrogen as substitutional nitrogen and suppress the incorporation of nitrogen as NV$^-$. The NV$^-$ background of the surface normalized deposition series was too high for laser writing. This necessitates the exploration of reactor parameters with higher sensitivity.

Lastly, the deposition film was PCD diamond for four of the eight deposition in the second experimental series with surface normalization. PCD formed on the depositions with reactor pressure at 250 mbar and 370 mbar for substrate surface offcut of greater than 0.5$^\circ$ with respect to the (100) plane. The high nitrogen concentration in the gas phase increases growth rate. The high growth rate can lead to PCD due to the rapid growth rate of nucleation sites. As PCD is not ideal for quantum technology, nitrogen gas phase composition should be reduced to reduce growth rate and suppress the deposition of PCD.

The resolution of quantum sensors with NV$^-$ defects can be improved through addressing materials engineering challenges through creation of ideal substrates for laser writing. The characteristics of a laser writing substrate include a delta doped layer of single substitutional nitrogen. The manufacturing of these substrates requires precise control of gas phase composition to minimize transition layer thickness of a delta doped region. Further control of nitrogen defect selectivity within the delta doped layer is necessary to reduce NV$^-$ background and incorporation single substitutional nitrogen.

\section{Proposed Future Work}
\indent
 
The conclusions of this work state the current CVD reactor, DS4, is not suitable for creating delta doped layers due to long residence times, high growth rate, and unitentional nitrogen concentration. A different reactor design, such as the NIRIM style reactor, would be suited for delta doping due to the laminar flow rate of deposition gases. 

Reactor pressure and substrate offcut exhibited low sensitivity with respect to nitrogen defect selection at high gas phase nitrogen concentration. Exploration of other deposition parameters with higher sensitivity. Potential deposition parameters include: low gas phase nitrogen concentration or different nitrogen sources.

In addition to nitrogen related defect selectivity, deposition parameters that produce SCD as opposed to PCD must be determined. Lower nitrogen gas phase composition would reduce the growth rate. Lower growth rates would reduce the growth rate of adatom nucleation at higher energy configurations, i.e. on plane as opposed to step edges, and thus reduce PCD depositions \cite{VanDerDrift1967EVOLUTIONARYLAYERS}.

Lastly, accurate characterization of nitrogen related defects through FTIR is valuable to non-destructively determine concentration of single substitutional nitrogen. The proposed method was validated using EPR spectroscopy. The proposed method deconvolutes a complicated absorption band in FITR using seven Gaussian peaks. To ensure the seven nitrogen related peaks used for the fit are present, FTIR data can be take at low temperatures to reduce the thermal broadening of the peaks to allow each peak to be more easily resolved. For this, a low temperature attachment for the FTIR is necessary.

This continued work would improve the quality of single crystal diamond laser writing substrates with a delta doped substitutional nitrogen layer. The improvement of the laser writing substrate will improve the resolution of quantum sensors made of NV$^-$ defects.