\chapter{Low Methane Growth Regime in High Growth Rate Reactor}
\label{chLowCH4}
\indent

This chapter presents the results of a deposition series grown at low methane concentrations in a high power, high growth rate chemical vapor deposition (CVD) diamond reactor. These experiments were designed to assess the quality of growth under not optimized, low growth rate, operation region of the high power, high growth rate of Diamond System 4 (DS4). As previously established, CVD diamond growth occurs in a thermodynamically metastable regime, where even subtle variations in substrate surface condition can influence growth quality. While Chapter \ref{chSurfacePrep} determined etching methods to normalize surface conditions for high quality depositions, this chapter will ensure high quality single crystal diamond growth occurs in the non-ideal operation region of DS4.

In contrast to the reactor's previous optimization for high growth rates, the present study explores a low methane concentration regime that results in reduced growth rates and deviates from standard operating conditions. Establishing reliable growth in this regime is critical for enabling precise control over nitrogen dopants in layered structures. In particular, slowing the growth rate provides a pathway to minimize line widths of doped layers through near halted growth rates during introduction and cessation of nitrogen gas flow into the reactor. Thus, this work is motivated not only by understanding growth behavior under not optimized condition but also by the need to maintain crystalline quality during layer device fabrication.

A key outcome of this study is the presence of nitrogen incorporation despite the absence of nitrogen in the feed gas. This unintentional nitrogen doping is a critical observation because it established a baseline nitrogen incorporation for this reactor. In response to the unintended nitrogen incorporation, a series of post-processing treatments were implemented. These processes and their resulting effects are discussed in details.

\section{Low Methane Growth in DS4}
\indent

Prior to the beginning of this dissertation work, Yajun Gu, a doctoral graduate student researcher in the MSU-Fraunnhofer lab, had optimized high quality single  crystal diamond growth on DS4, a MPACVD reactor that was described in Section \ref{DS4}. The optimization of DS4 included the effects of methane concentration and reactor pressure on growth rate and compared these results to other bell jar style reactors \cite{Gu2012}. The study of methane concentration on growth rate evaluated 3 percent, 5 percent, 7 percent, and 9 percent methane. At 3 percent methane, the growth rate measured approximately 18 $\mu m$ per hour. 

\section{Methane Concentration Effects on Growth Rate}
\section{Formation of Unepitaxial Crystallites}
\section{Unitentional Nitrogen Incorporation}
\section{Post-Processing to Reduce Nitrogen Vacancy Center}
\subsection{High Pressure Annealing}
\subsection{High Temperature Hydrogen Environment to Passivate Nitrogen Vacancy Center}
